---
title: "(Ficitonal) Perinatal HIV Data"
author: "Elliott Brannon"
date: "Tuesday, January 27, 2015"
output: html_document
runtime: shiny
---

Note: All data are fictional.

This is an interactive document describing fictional perinatal HIV. Although the data are fictional, it could be used as a template to create an easily updatable Annual Report. The line list of this fictional data can be downloaded here:
[data](data.Rmd)

## Exposures, Perinatal HIV Transmissions and Transmission Rate
When switching from plot to table you may need to reselect the variable or year to refresh the plot/table.

```{r echo=FALSE, warning=FALSE}
library(ggplot2)
library(reshape2)
set.seed(1234)




for (byear in 2000:2013) {
  n <- sample(150:205, 1)
  count <- 1
  hiv <- rbinom(n,1,.2)
  pncnum <- sample(0:22, n, replace=TRUE)
  black <- rbinom(n, 1, .8)
  white <- rbinom(n, 1, .2)
  race <- sample(c("Black","White","Hispanic","Multiple"), n, replace=TRUE)
  mage <- sample(15:40, n, replace=TRUE)
  region <- sample(1:9, n, replace=TRUE)
  trans <- sample(1:3,n, replace=TRUE)
  timely <- rbinom(n, 1, .1)
  arvpreg <- rbinom(n, 1, .2)
  arvlabor <- rbinom(n, 1, .1)
  cesarean <- rbinom(n, 1, .1)
  arvneo <- rbinom(n, 1, .05)
  breastfed <- rbinom(n, 1, .1)
  timing <- sample(0:700, n, replace=TRUE)
  vl <- sample(0:10000, n, replace=TRUE)
  birthwt <- sample(1000:5000, n, replace=TRUE)
  gestage <- sample(25:42, n, replace=TRUE)
  fips <- sample(1:64,n,1)*2 - 1 + 22000
  status <-(rbinom(n,1,.9)+1*hiv)

  data <- data.frame(1:n, byear,count, hiv, pncnum, black, white, mage, region, trans, timely, arvpreg, arvlabor, cesarean,
                     arvneo, breastfed, timing, vl, birthwt, gestage, fips, status)

  if(byear == 2000){ peds <- data.frame(data)
  } else {peds <- data.frame(rbind(peds, data))}
  
}


pedsbyyear <- aggregate(peds, by=list(peds$byear), FUN=sum)




pedsbyyear$rate <- pedsbyyear$hiv/pedsbyyear$count

inputPanel(
  selectInput("var1", label = "Choose Variable:",
              choices = c("exposure","hiv","rate")),
  selectInput("year1", label = "Start Year:",
              choices = c(2000:2013)),
  selectInput("year2", label = "End Year:",
              choices = c(2000:2013),selected=2013)  
  
)





  
tabsetPanel(
      tabPanel('plot',

renderPlot({
  x <- (as.numeric(input$year1)-1999):(as.numeric(input$year2)-1999)
  years <- as.numeric(input$year1):as.numeric(input$year2)
  data1 <- switch(input$var1, 
        "exposure" = pedsbyyear$count[x],
        "hiv" = pedsbyyear$hiv[x],
         "rate" = pedsbyyear$rate[x])
 plot(data1  ~ years, ylab=input$var1)
 
 }),
 
renderPlot({
  x <- (as.numeric(input$year1)-1999):(as.numeric(input$year2)-1999)
  years <- as.numeric(input$year1):as.numeric(input$year2)
 
 ## Set up data
line.x <- years
bar.x <- years
bar.y <- pedsbyyear$count[x]
line.x <- years
line.y <- pedsbyyear$rate[x]
x.range <- range(bar.x, line.x)


## Plot the data
par(mar=c(5,4,4,4)+0.1)    ## Make enough room for both labels
plot(y=bar.y, x=bar.x, type='h', lwd=25, lend=4, xlim=x.range, xlab="Year", ylab="", main="Perinatal HIV Transmissions and Exposure")
par(new=TRUE)
plot(y=line.y, x=line.x, col='red', type='l', xlim=x.range, axes=FALSE, xlab='', ylab='')

## Set up the axes and labels
yLabels <- seq(0, 1, .02)
axis(side=4, at=yLabels, labels=sprintf(round(100*yLabels), fmt="%2.f%%"), las=1, col="red", col.axis="red")
#at = axTicks(4)
#mtext(side = 4, text = at, at = at, col = "red", line = 1) 
## Add x-axis labels; this allows customization of the how far out labels are
mtext(text='Perinatal HIV Transmission Rate', side=4, line=3, col='red')
mtext(text='Perinatal HIV Exposures', side=2, line=2)
 
 })
),


tabPanel('table', renderDataTable({
  x <- (as.numeric(input$year1)-1999):(as.numeric(input$year2)-1999)
  years <- as.numeric(input$year1):as.numeric(input$year2)
  temp <- subset(pedsbyyear, Group.1>=input$year1 & Group.1<=input$year2)
  temp <- temp[c("Group.1","hiv","count", "rate")]
  colnames(temp) = c("byear","hiv","exposure","rate")
  temp}, options = list(paging = FALSE, searching=FALSE)))
  
)



```

# All variables

When switching from plot to table you may need to reselect the variable or year to refresh the plot/table.

```{r echo=FALSE}


pie2= function (x, labels = names(x), edges = 200, radius = 0.8, clockwise = FALSE, 
          init.angle = if (clockwise) 90 else 0, density = NULL, angle = 45, 
          col = NULL, border = NULL, lty = NULL, tcklen=1, main = NULL, ...) 
{
  if (!is.numeric(x) || any(is.na(x) | x < 0)) 
    stop("'x' values must be positive.")
  if (is.null(labels)) 
    labels <- as.character(seq_along(x))
  else labels <- as.graphicsAnnot(labels)
  x <- c(0, cumsum(x)/sum(x))
  dx <- diff(x)
  nx <- length(dx)
  plot.new()
  pin <- par("pin")
  xlim <- ylim <- c(-1, 1)
  if (pin[1L] > pin[2L]) 
    xlim <- (pin[1L]/pin[2L]) * xlim
  else ylim <- (pin[2L]/pin[1L]) * ylim
  dev.hold()
  on.exit(dev.flush())
  plot.window(xlim, ylim, "", asp = 1)
  if (is.null(col)) 
    col <- if (is.null(density)) 
      c("white", "lightblue", "mistyrose", "lightcyan", 
        "lavender", "cornsilk")
  else par("fg")
  if (!is.null(col)) 
    col <- rep_len(col, nx)
  if (!is.null(border)) 
    border <- rep_len(border, nx)
  if (!is.null(lty)) 
    lty <- rep_len(lty, nx)
  angle <- rep(angle, nx)
  if (!is.null(density)) 
    density <- rep_len(density, nx)
  twopi <- if (clockwise) 
    -2 * pi
  else 2 * pi
  t2xy <- function(t) {
    t2p <- twopi * t + init.angle * pi/180
    list(x = radius * cos(t2p), y = radius * sin(t2p))
  }
  for (i in 1L:nx) {
    n <- max(2, floor(edges * dx[i]))
    P <- t2xy(seq.int(x[i], x[i + 1], length.out = n))
    polygon(c(P$x, 0), c(P$y, 0), density = density[i], angle = angle[i], 
            border = border[i], col = col[i], lty = lty[i])
    P <- t2xy(mean(x[i + 0:1]))
    lab <- as.character(labels[i])
    if (!is.na(lab) && nzchar(lab)) {
      lines(c(1, tcklen) * P$x, c(1, tcklen) * P$y)
      text((tcklen+.1) * P$x, (tcklen+.1) * P$y, labels[i], xpd = TRUE, 
           adj = ifelse(P$x < 0, 1, 0), ...)
    }
  }
  title(main = main, ...)
  invisible(NULL)
}

temptable=function(year, variable){
    temp <- peds[c("count","byear",variable)]
  temp <- aggregate(temp$count, by=list(temp$byear, temp[,3]), FUN=sum)
  colnames(temp) <- c("byear",variable, "count")
  temp <- subset(temp, byear == year & temp[,2] != '')
  temp$percent <- round(temp$count/sum(temp$count)*100)
 temp
}

piechart2=function(year, variable){
  temp <- peds[c("count","byear",variable)]
  temp <- aggregate(temp$count, by=list(temp$byear, temp[,3]), FUN=sum)
  colnames(temp) <- c("byear",variable, "count")
  temp <- subset(temp, byear == year & temp[,2] != '')
  par(mar=c(0,0,0,0))
  par(cex=1.2, pin=c(4,4))
  pct <- round(temp$count/sum(temp$count)*100)
  lbls <- temp[,2]
  lbls <- paste(lbls,"\n N = ", temp$count," (", sep="") # add percents to labels 
  lbls <- paste(lbls,pct, "%)",sep="")
  pie2(temp$count, labels=lbls, main=paste("Pie chart of",variable, "in", year), tcklen=1.2, line=2)
}

shinyUI(fluidPage(
  titlePanel("All variables"),
sidebarLayout(
sidebarPanel(
inputPanel(
  selectInput("year", label = "Year:",
              choices = c(2000:2013)),
  selectInput("choice", label = "Variable:",
              choices = c("hiv" 
,"pncnum" 
,"black" 
,"white"
,"mage" 
,"region" 
,"trans"
,"timely"
,"arvpreg"
,"arvlabor"
,"cesarean"
,"arvneo"
,"breastfed"
,"timing"
,"vl"
,"birthwt"
,"gestage"
,"status")
  )
)),

mainPanel(
  
  tabsetPanel(
      tabPanel('plot',
               renderPlot({piechart2(as.numeric(input$year), input$choice)}, height = 600, width = 600)),
      tabPanel('table', renderDataTable({ temptable(input$year, input$choice) }, options = list(paging = FALSE, searching=FALSE))
  
)

)))))



```


## All variables 2

```{r echo=FALSE, warning=FALSE}

yearvar <- function(year=2000, var="region") {
  temp <- subset(peds[c("byear","count",var)], byear == year)
  temp <- aggregate(temp$count, by=list(temp$byear, temp[,3]), FUN=sum)
  colnames(temp) <- c("byear",var, "count")
  temp$pct <- round((temp$count/sum(temp$count))*100,0)
  temp
}

yearsvar <- function(year1=2000, year2=2013, var="region") {
  #temp <- subset(peds[c("byear","count",var)], eval(parse(text=var)) == var2)
  temp <- peds[c("byear","count",var)]
  temp <- aggregate(temp$count, by=list(temp$byear, temp[,3]), FUN=sum)
  colnames(temp) <- c("byear",var, "count")
  temp
}

bar <- function(year = 2000, var = "region", orderby = 0) {
  temp <- yearvar(year=year, var=var)
    if(orderby == 1){
      temp[var] <- factor(temp[,var], levels = temp[,var][order(temp["count"])])
    } else {
          temp[var] <- factor(temp[,var], levels = temp[,var][order(-temp[var])])    
    }
  var <- paste("factor(",var,")",sep="" )
  temp$labels <- paste(temp$count, ", ", temp$pct, "%", sep="")
  ggplot(temp, aes_string(x=var, y="count")) + 
    geom_bar(stat="identity") + coord_flip() + 
    geom_text(aes_string(x=var, y="count", label="labels"))
}

line <- function(year1 =2000, year2 = 2013, var="region"){
  temp <- yearsvar(year1 = 2000, year2 = 2013, var=var)
  ggplot(temp, aes_string(x="byear", y="count", group=var, color=var)) + 
    geom_line(stat="identity") +  
    geom_text(aes_string(x="byear", y="count", label="count"))  
}



shinyUI(fluidPage(
  titlePanel("All variables"),
sidebarLayout(
sidebarPanel(
inputPanel(
  selectInput("year_", label = "Year:",
              choices = c(2000:2013)),
  selectInput("choice_", label = "Variable:",
              choices = c("hiv" 
,"pncnum" 
,"black" 
,"white"
,"mage" 
,"region" 
,"trans"
,"timely"
,"arvpreg"
,"arvlabor"
,"cesarean"
,"arvneo"
,"breastfed"
,"timing"
,"vl"
,"birthwt"
,"gestage"
,"status")
  ),
 selectInput("order", label = "Order by:",
              choices = c("name"= 0, "count" = 1))

)),

mainPanel(
  


               renderPlot({bar(year=input$year_, var=input$choice_, order=input$order)}, height = 600, width = 600))
  
)

))




```




```{r echo=FALSE}

library(maps)
data(county.fips)
county.fips <- subset(county.fips, fips>22000 & fips<23000)
colors = c("#F1EEF6", "#D4B9DA", "#C994C7", "#DF65B0", "#DD1C77", 
    "#980043")

shinyUI(fluidPage(
  titlePanel("Map of Perinatal HIV exposures"),
sidebarLayout(
sidebarPanel(
inputPanel(
  selectInput("year__", label = "Year:",
              choices = c(2000:2013))
)),


mainPanel(

               renderPlot({


peds_cnty <- yearvar(year = input$year__, var="fips")
peds_cnty$colorBuckets <- peds_cnty$count

peds_cnty$colorBuckets <- as.numeric(cut(peds_cnty$count, c(0, 2, 4, 6, 8, 
    10, 100)))
colorsmatched <- peds_cnty$colorBuckets[match(as.numeric(county.fips$fips), peds_cnty$fips)]


map("county", col = colors[colorsmatched], fill = TRUE, region = "louisiana", myborder=c(.1,.1))

title(paste("Perinatal HIV exposures,", input$year__), line =2)

leg.txt <- c("1 - 2", "3 - 4", "5 - 6", " 7 - 8", "9 - 10")
legend("bottomright", leg.txt, horiz = TRUE, fill = colors)
},  height = 600, width = 600
)))))



```

## Maternal Age Category

```{r echo=FALSE, warning=FALSE}

peds$magecat <- ifelse(peds$mage <= 19, '15 - 19', ifelse(peds$mage <= 24, '20 - 24', 
                ifelse(peds$mage <= 34, '25 - 34',ifelse(peds$mage <= 44, '35 - 44', '>45'))))
mage <- aggregate(peds$count, by=list(peds$byear, peds$magecat), FUN=sum)
colnames(mage) <- c("byear","magecat","count")


peds$pnccat <- ifelse(peds$pncnum <= 1, 'No PNC', ifelse(peds$pncnum <= 4, '1 - 4 visits', 
                ifelse(peds$pncnum <= 10, '5 - 10 visits',ifelse(peds$pncnum <= 16, '11 - 16 visits', '17+ visits'))))
pnc <- aggregate(peds$count, by=list(peds$byear, peds$pnccat), FUN=sum)
colnames(pnc) <- c("byear","pnccat","count")

piechart=function(type, year){

  type2 <- paste("factor(",type,"cat)",sep="")

  test <- subset(eval(parse(text=type)), byear == year)
        
  
  #test <- test %>% group_by(byear) %>% mutate(pos = cumsum(count)- count/2)
  test$pos<-(pos<-cumsum(test$count)-diff(c(0,cumsum(test$count)))*(1-0.5)) 
  
  ggplot(data=test, aes_string(x="''", y="count", fill=type2)) +
  geom_bar(stat="identity", width=1) +
  geom_text(aes(x=1.75,y=pos, label = scales::percent((count)/sum(count))), size=8) +  # note y = pos
  labs(title = year ) +
  coord_polar(theta = "y") +
    theme(
      plot.title = element_text(size = rel(2))
      ,plot.background = element_blank()
      ,panel.grid.major = element_blank()
      ,panel.grid.minor = element_blank()
      ,panel.border = element_blank()
      ,axis.text.y = element_blank()
      ,axis.text.x = element_blank()
      ,axis.title.x=element_blank()
      ,axis.title.y=element_blank()
      ,axis.ticks=element_blank()
    )
}

inputPanel(
  selectInput("year_agecat", label = "Year:",
              choices = c(2000:2013)))

renderPlot({
piechart("mage", as.numeric(input$year_agecat))

})

renderPlot({
    test <- subset(mage, byear == input$year_agecat)
    pct <- round(test$count/sum(test$count)*100)
    lbls <- test$magecat
lbls <- paste(lbls,"\n N = ", test$count," (", pct, sep="") # add percents to labels 
lbls <- paste(lbls,"%)",sep="")
pie(test$count, labels=lbls)

})


```

#Number of Prenatal Care Visits

```{r echo=FALSE}

inputPanel(
  selectInput("year_pnccat", label = "Year:",
              choices = c(2000:2013)))

renderPlot({
piechart("pnc", as.numeric(input$year_pnccat))
})

```



## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r, echo=FALSE, results=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package="shiny"),
  options=list(
    width="100%", height=550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.

http://esbrannon.shinyapps.io/PerinatalHIV/test.Rmd


